{\fontfamily{ptm}\selectfont A Parameter Free Accelerated Proximal Gradient Method Without Restarting }

### 2025-01-23 Notebook LaTEX/R-WAPG/paper3.tex
>w<, changes

        It has the ability to upper bound $F(x_k) - F(x^*)$ for sequences $(t_k)_{k \ge 0}$ that follows a rule much weaker than Nesterov's update rule. 
        In addition to a convergence claim of $F(x_k)- F(x^*)$ for a much more flexible choice of $(t_k)_{k \ge 1}$. 
        Secondly, we present an alternative to restarting that performs well empirically inspired by a small detail in the convergence proof of R-WAPG. 
        It also has descriptive power to describe several variants of FISTA in the literatures.  

@@

@@

        Our practical contribution is an algorithm inspired by a detail in our convergence proof which we call it ``Parameter Free R-WAPG'' (See Algorithm \ref{alg:free-rwapg}). 

@@

        \begin{definition}[R-WAPG intermediate form]\label{def:r-wapg-intermediate}\;\\
            Assume $\mu < L$ and let $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ given by Definition \ref{def:rwapg-seq}. 
            Initialize any $x_1, v_1$ in $\RR^n$. 
            For $k \ge 1$, the algorithm generates sequence of vector iterates $(y_{k}, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
            \begin{tcolorbox}
                For $k = 1, 2, \ldots$
                \begin{align*} 
                    & y_{k} = 
                    \left(
                        1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                    \right)^{-1}
                    \left(
                        v_{k + 1} + 
                        \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                    \right), 
                    \\
                    & x_{k + 1} = 
                    y_k - L^{-1} \mathcal G_L y_k, 
                    \\
                    & v_{k + 1} = 
                    \left(
                        1 + \frac{\mu}{L \alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k + 
                        \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                    \right) - \frac{1}{L\alpha_{k}}\mathcal G_L y_k. 
                \end{align*}
            \end{tcolorbox}
        \end{definition}
        \begin{remark}
            This form of APG is rarely identified in the literatures. 
            The closest algorithm that fits the form but with $\mu = 0$ is Chapter 12 of in Ryu and Yin's Book \cite{ryu_large-scale_2022}, right after Theorem 17. 
            We created this form which makes the math that follows simpler. 
            The inspiration of using this as an intermediate representation was inspired by solving Exercise 12.1 in the same Ryu and Yin's Book. 
        \end{remark}
        \begin{definition}[R-WAPG similar triangle form]\label{def:r-wapg-st-form} \; \\
            Given any $(x_1, v_1)$ in $\RR^n$. 
            Assume $\mu < L$.
            Let the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k\ge 0}$ be given by Definition \ref{def:rwapg-seq}. 
            For $k \ge 1$, the algorithm generates sequences of vector iterates $(y_k, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
            \begin{tcolorbox}
                For $k=1, 2, \ldots $
                \begin{align*}
                    & y_k = 
                    \left(
                        1 + \frac{L - L\alpha_k}{L\alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k + 
                        \left(\frac{L - L\alpha_k}{L\alpha_k - \mu} \right) x_k
                    \right), 
                    \\
                    & x_{k + 1} = 
                    y_k - L^{-1} \mathcal G_L y_k, 
                    \\
                    & v_{k + 1} = 
                    x_{k + 1} + (\alpha_k^{-1} -1)(x_{k + 1} - x_k). 
                \end{align*}    
            \end{tcolorbox}
        \end{definition}
        \begin{remark}
            The word similar triangle form can be traced back to several literatures. 
            The term ``Method of Similar Triangle" was used for Algorithm (6.1.19) in Nesterov's book \cite{nesterov_lectures_2018}, but without the necessary graphical illustrations to clarify it. 
            Equation (2), (3), (4) in \cite{chambolle_convergence_2015} is a similar triangle formulation of FISTA with $\mu = 0$. 
            To see graphical visualization on why such term is used to describe the APG algorithm in the literatures, see 
            (3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022}.

        \end{remark}
        \begin{definition}[R-WAPG momentum form]\label{def:r-wapg-momentum-form}
            Given any $y_1 = x_1 \in \RR^n$, and sequences $(\rho_k)_{k \ge 0}, (\alpha_k)_{k\ge 0}$ Definition \ref{def:rwapg-seq}. 
            The algorithm generates iterates $x_{k + 1}, y_{k + 1}$ For $k = 1, 2, \cdots $ by the procedures: 
            \begin{tcolorbox}
                For $k=1, 2,\ldots $
                \begin{align*}
                    & x_{k + 1} = y_k - L^{-1}\mathcal G_Ly_k, 
                    \\
                    & 
                    y_{k + 1} = 
                    x_{k + 1} + 
                    \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k). 
                \end{align*}    
            \end{tcolorbox}
            In the special case where $\mu = 0$, the momentum term can be represented without relaxation parameter $\rho_k$: 
            $$
                (\forall k \ge 1)\quad \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}} 
                = \alpha_{k + 1}(\alpha_k^{-1} - 1). 
            $$
        \end{definition}
        \begin{remark}
            This format fits with (2.2.19) in Nesterov's book \cite{nesterov_lectures_2018}, however, the sequence $(\alpha_k)_{k \ge 0}$ would be given by a different rule. 
            See Theorem \ref{thm:r-wapg-on-cham-doss} and Lemma \ref{lemma:inverted-fista-seq} to see a specific choice of $(\alpha_k)_{k \ge0}, (\rho_k)_{ k\ge 0}$ such this equivalent form of R-WAPG is in fact two possible variants of the FISTA algorithm.
        \end{remark}

>>>>

@@

        It can upper bound $F(x_k) - F(x^*)$ for sequences $(t_k)_{k \ge 0}$ that doesn't follow Nesterov's update rule and it unifies the convergence results of several Euclidean variants of FISTA. 
        Secondly, we present an alternative to restarting that performs well empirically inspired by a detail in the convergence proof of R-WAPG with extensive numerical experiments. 

@@

        Our practical contribution is an algorithm inspired by a detail in the proof of Proposition \ref{prop:wagp-convergence} which is called ``Parameter Free R-WAPG'' (See Algorithm \ref{alg:free-rwapg}). 

@@

        \begin{definition}[R-WAPG intermediate form]\label{def:r-wapg-intermediate}\;\\
            Assume $\mu < L$ and let $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ given by Definition \ref{def:rwapg-seq}. 
            Initialize any $x_1, v_1$ in $\RR^n$. 
            For $k \ge 1$, the algorithm generates sequence of vector iterates $(y_{k}, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
                For $k = 1, 2, \ldots$
                    & y_{k} = 
                        v_{k + 1} + 
        \end{definition}
            This form of APG is rarely identified in the literatures. 
            The closest algorithm that fits the form but with $\mu = 0$ is Chapter 12 of in Ryu and Yin's Book \cite{ryu_large-scale_2022}, right after Theorem 17. 
            We created this form which makes the math that follows simpler. 
            The inspiration of using this as an intermediate representation was inspired by solving Exercise 12.1 in the same Ryu and Yin's Book. 
        \begin{definition}[R-WAPG similar triangle form]\label{def:r-wapg-st-form} \; \\
            Given any $(x_1, v_1)$ in $\RR^n$. 
            Assume $\mu < L$.
            Let the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k\ge 0}$ be given by Definition \ref{def:rwapg-seq}. 
            For $k \ge 1$, the algorithm generates sequences of vector iterates $(y_k, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
                For $k=1, 2, \ldots $
                    & y_k = 
                    x_{k + 1} + (\alpha_k^{-1} -1)(x_{k + 1} - x_k). 
        \end{definition}
            The word similar triangle form can be traced back to several literatures. 
            The term ``Method of Similar Triangle" was used for Algorithm (6.1.19) in Nesterov's book \cite{nesterov_lectures_2018}, but without the necessary graphical illustrations to clarify it. 
            Equation (2), (3), (4) in \cite{chambolle_convergence_2015} is a similar triangle formulation of FISTA with $\mu = 0$. 
            To see graphical visualization on why such term is used to describe the APG algorithm in the literatures, see 
            (3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022}.
        \begin{definition}[R-WAPG momentum form]\label{def:r-wapg-momentum-form}
            Given any $y_1 = x_1 \in \RR^n$, and sequences $(\rho_k)_{k \ge 0}, (\alpha_k)_{k\ge 0}$ Definition \ref{def:rwapg-seq}. 
            The algorithm generates iterates $x_{k + 1}, y_{k + 1}$ For $k = 1, 2, \cdots $ by the procedures: 
                For $k=1, 2,\ldots $
                    & x_{k + 1} = y_k - L^{-1}\mathcal G_Ly_k, 
                    y_{k + 1} = 
                    \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k). 
            In the special case where $\mu = 0$, the momentum term can be represented without relaxation parameter $\rho_k$: 
                (\forall k \ge 1)\quad \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}} 
                = \alpha_{k + 1}(\alpha_k^{-1} - 1). 
        \end{definition}
            This format fits with (2.2.19) in Nesterov's book \cite{nesterov_lectures_2018}, however, the sequence $(\alpha_k)_{k \ge 0}$ would be given by a different rule. 
            See Theorem \ref{thm:r-wapg-on-cham-doss} and Lemma \ref{lemma:inverted-fista-seq} to see a specific choice of $(\alpha_k)_{k \ge0}, (\rho_k)_{ k\ge 0}$ such this equivalent form of R-WAPG is in fact two possible variants of the FISTA algorithm.
